---
title: "multi-tissue analysis (LD merge =T)"
author: "XSun"
date: "2023-12-13"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

For each trait, we sorted the tissues by PVE explained by genes, then the top5 un-correlated tissues were selected for the multi-tissue analysis.

[link to deciding tissues](https://xsun1229.github.io/ctwas_Tibetan/uncorrelated_tissue.html)

# Data 

## Traits and selected tissues


| Trait | Tissue                 | PVE_gene            |
|-------|------------------------|---------------------|
| LB    | Lung                   | 0.0410350101026084  |
|       | Whole_Blood            | 0.0110875627240866  |
|       | Heart_Left_Ventricle   | 0.0107697033082494  |
|       | Artery_Coronary        | 0.00930578096193273 |
|       | Kidney_Cortex          | 0.00426258259876705 |
| MC    | Liver                  | 0.0178648723821638  |
|       | Artery_Tibial          | 0.0134706822007496  |
|       | Heart_Atrial_Appendage | 0.0100733429644444  |
|       | Pancreas               | 0.00867082220578753 |
|       | Lung                   | 0.0085924869917463  |
| OxHb  | Heart_Left_Ventricle   | 0.0633139032762666  |
|       | Kidney_Cortex          | 0.0397523122784917  |
|       | Liver                  | 0.0231916966396772  |
|       | Artery_Coronary        | 0.0135241915935851  |
|       | Lung                   | 0.0104129270975456  |
| Pulse | Heart_Atrial_Appendage | 0.0492054125707622  |
|       | Kidney_Cortex          | 0.0402428509363196  |
|       | Artery_Tibial          | 0.0173924354650824  |
|       | Whole_Blood            | 0.0148621621982472  |
|       | Liver                  | 0.00959987160023217 |



## Functions used

```{r }
suppressMessages(library(ctwas))
suppressMessages(library(data.table))
suppressMessages(library(RSQLite))
suppressMessages(library(DT))
suppressMessages(library(tools))
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))

thin <- 0.1

process_data <- function(weight, outputdir, outname, z_snp, z_gene, display_datatable = TRUE) {
  # Splitting and constructing weight file paths
  weight <- unlist(strsplit(weight, split=","))
  weight <- paste0("/project2/xinhe/shared_data/multigroup_ctwas/weights/mashr_models/mashr_", weight, ".db")

  # Reading cTWAS results
  ctwas_res <- fread(paste0(outputdir, outname, ".susieIrss.txt"))

  # Initialize gene_info dataframe
  gene_info <- data.frame(gene=character(), genename=character(), gene_type=character(), weight=character())
  for (i in seq_along(weight)) {
    sqlite <- dbDriver("SQLite")
    db <- dbConnect(sqlite, weight[i])
    query <- function(...) dbGetQuery(db, ...)
    gene_info_current <- query("SELECT gene, genename, gene_type FROM extra")
    dbDisconnect(db)

    gene_info_current$weight <- weight[i]
    gene_info <- rbind(gene_info, gene_info_current)
  }

  # Processing gene_info
  gene_info$weight <- sapply(gene_info$weight, function(x) rev(unlist(strsplit(file_path_sans_ext(x), "/")))[1])
  gene_info$id <- paste(gene_info$gene, gene_info$weight, sep="|")

  # Adding gene names to cTWAS results
  ctwas_res$genename[ctwas_res$type != "SNP"] <- gene_info$genename[match(ctwas_res$id[ctwas_res$type != "SNP"], gene_info$id)]

  # Adding z values
  ctwas_res$z[ctwas_res$type == "SNP"] <- z_snp$z[match(ctwas_res$id[ctwas_res$type == "SNP"], z_snp$id)]
  ctwas_res$z[ctwas_res$type != "SNP"] <- z_gene$z[match(ctwas_res$id[ctwas_res$type != "SNP"], z_gene$id)]

  # Sorting and displaying datatable if required
  ctwas_res <- ctwas_res[order(-ctwas_res$susie_pip),]
  if (display_datatable) {
    datatable(ctwas_res[ctwas_res$type != "SNP" & ctwas_res$susie_pip > 0.8,])
  }
  
  return(ctwas_res)
}

process_gene_data <- function(ctwas_res, weight = NULL) {
  # Aggregate susie_pip
  df_gene <- aggregate(ctwas_res$susie_pip[ctwas_res$type != "SNP"], 
                       by = list(ctwas_res$genename[ctwas_res$type != "SNP"]), 
                       FUN = sum)
  colnames(df_gene) <- c("genename", "combined_pip")

  # Optional part for dropping duplicated gene names
  # Uncomment this part if needed
  # df_gene <- df_gene[!(df_gene$genename %in% names(which(table(ctwas_res$genename) > length(weight)))),]

  # Collect tissue-level results
  all_tissue_names <- unique(ctwas_res$type[ctwas_res$type != "SNP"])

  df_gene_pips <- matrix(NA, nrow = nrow(df_gene), ncol = length(all_tissue_names))
  colnames(df_gene_pips) <- all_tissue_names

  ctwas_gene_res <- ctwas_res[ctwas_res$type != "SNP",]

  for (i in 1:nrow(df_gene)) {
    gene <- df_gene$genename[i]
    ctwas_gene_res_subset <- ctwas_gene_res[ctwas_gene_res$genename == gene,]
    df_gene_pips[i, ctwas_gene_res_subset$type] <- ctwas_gene_res_subset$susie_pip
  }

  df_gene <- cbind(df_gene, df_gene_pips)

  # Sort by combined PIP
  df_gene <- df_gene[order(-df_gene$combined_pip),]
  df_gene <- df_gene[, apply(df_gene, 2, function(x) {!all(is.na(x))})]  # Drop genes not imputed in any tissue

  # Determine number of rows to display
  n_rows <- max(sum(df_gene$combined_pip > 0.8), 20)

  # Select top rows based on the criteria
  df_to_display <- head(df_gene, n_rows)
  
  # Return a datatable
 # return(DT::datatable(df_to_display,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Genes with combined pip > 0.8, 0 means pip=0; NA means no such gene in this tissue'),options = list(pageLength = 10) ))
    datatable_display <- DT::datatable(df_to_display, 
                                     caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black; font-size:150%;', 
                                     'Genes with combined pip > 0.8, 0 means pip=0; NA means no such gene in this tissue'), 
                                     options = list(pageLength = 10))
  return(df_to_display)
}

process_ctwas_results <- function(ctwas_parameters, ctwas_res) {
    # Process parameters
    para <- cbind(ctwas_parameters$group_size,
                  ctwas_parameters$group_prior,
                  ctwas_parameters$group_prior_var,
                  c(NA, ctwas_parameters$enrichment),
                  ctwas_parameters$group_pve)
    para <- as.data.frame(cbind(rownames(para), para))
    colnames(para) <- c("group","group_size","group_prior","group_prior_var","enrichment","group_pve")

    # Filter ctwas results for genes and high PIP
    ctwas_res_gene <- ctwas_res[ctwas_res$type != "SNP",]
    
    max_susie_pip_by_type <- ctwas_res_gene %>%
    group_by(type) %>%
    summarize(max_susie_pip = max(susie_pip, na.rm = TRUE))

    # Merge summary with parameters
    summary <- merge(max_susie_pip_by_type, para, by.x = "type", by.y = "group", all.y = TRUE)

    return(summary)
}

attributable_pip <- function(df, combined_pip_col) {
  # Identify the relevant columns (those ending with '_nolnc')
  #relevant_cols <- names(df)[grepl("_nolnc$", names(df))]
  relevant_cols <- names(df)[!names(df) %in%c( "combined_pip","genename")]
  # Create and rearrange columns dynamically
  df <- df %>%
    mutate(across(all_of(relevant_cols), 
                  ~ ifelse(!is.na(.x) & !is.na(combined_pip), 
                           .x / combined_pip * 100, NA), 
                  .names = "{.col}_Attributable_PIP (%)"))%>%
    {
      # Building the new column order
      new_order <- unlist(lapply(relevant_cols, function(col) {
        c(col, paste0(col, "_Attributable_PIP (%)"))
      }))
      select(., all_of(c("genename", combined_pip_col, new_order)))
    }

  
}


plot_stacked_histogram <- function(df, genename_col, combined_pip_col, attributable_cols) {
  
  df[[genename_col]] <- factor(df[[genename_col]], levels = df[[genename_col]])
  
  long_df_attributable <- df %>%
    select(all_of(c(genename_col, combined_pip_col, attributable_cols))) %>%
    pivot_longer(
      cols = all_of(attributable_cols),
      names_to = "Tissue",
      values_to = "Proportion"
    ) %>%
    mutate(Tissue = gsub("_Attributable_PIP \\(%\\)", "", Tissue))  # Modify Tissue names
  
  
  # Calculate the absolute values for each segment
  long_df_attributable <- long_df_attributable %>%
    mutate(Absolute_Value = Proportion / 100 * long_df_attributable[[combined_pip_col]])
  
  # Create the plot
  p <- ggplot(long_df_attributable, aes_string(x = genename_col, y = "Absolute_Value", fill = "Tissue")) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    labs(title = "Stacked Histogram of Combined PIP with Tissue Contributions",
         x = "Gene Name",
         y = "Combined PIP",
         fill = "Tissue") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  return(p)
}

```

# LB

```{r fig.height=6, fig.width=12,message=FALSE, warning=FALSE}
outputdir <- "/project/xinhe/xsun/ctwas/2.TIBETN/output/multi/LB/"
outname <- "LB_ctwas"
gwas_n <- 966
weight <- "Lung,Whole_Blood,Heart_Left_Ventricle,Artery_Coronary,Kidney_Cortex"

load(paste0(outputdir,outname,"_z_snp.Rd"))
load(paste0(outputdir,outname,"_z_gene.Rd"))

ctwas_parameters <- ctwas:::ctwas_summarize_parameters(outputdir = outputdir,
                                               outname = outname,
                                               gwas_n = gwas_n,
                                               thin = thin)

ctwas_parameters$convergence_plot

ctwas_res <- process_data(weight = weight,outputdir = outputdir,outname = outname,z_snp = z_snp,z_gene = z_gene)
para <- process_ctwas_results(ctwas_parameters = ctwas_parameters,ctwas_res = ctwas_res)
DT::datatable(para,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Parameters'))

df_gene <- process_gene_data(ctwas_res = ctwas_res)

df_gene <- attributable_pip(df_gene,combined_pip_col = "combined_pip")

DT::datatable(df_gene,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','top20 Genes sorted bu susiepip, 0 means pip=0; NA means no such gene in this tissue'),options = list(pageLength = 10) )

attributable_columns <- colnames(df_gene)[seq(4,12,by=2)]

plot_stacked_histogram(df_gene, "genename", "combined_pip", attributable_columns)

table_show <- ctwas_res[ctwas_res$genename %in% df_gene$genename,]
table_show <- rbind(table_show,ctwas_res[ctwas_res$type!="SNP" & ctwas_res$susie_pip > 0.8,])
table_show <- table_show[!duplicated(table_show$id),]
DT::datatable(table_show,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Detailed table for genes showed above'),options = list(pageLength = 5))

```

# MC

```{r fig.height=6, fig.width=12,message=FALSE, warning=FALSE}
outputdir <- "/project/xinhe/xsun/ctwas/2.TIBETN/output/multi/MC/"
outname <- "MC_ctwas"
gwas_n <- 966
weight <- "Liver,Artery_Tibial,Heart_Atrial_Appendage,Pancreas,Lung"

load(paste0(outputdir,outname,"_z_snp.Rd"))
load(paste0(outputdir,outname,"_z_gene.Rd"))

ctwas_parameters <- ctwas:::ctwas_summarize_parameters(outputdir = outputdir,
                                               outname = outname,
                                               gwas_n = gwas_n,
                                               thin = thin)

ctwas_parameters$convergence_plot

ctwas_res <- process_data(weight = weight,outputdir = outputdir,outname = outname,z_snp = z_snp,z_gene = z_gene)
para <- process_ctwas_results(ctwas_parameters = ctwas_parameters,ctwas_res = ctwas_res)
DT::datatable(para,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Parameters'))

df_gene <- process_gene_data(ctwas_res = ctwas_res)

df_gene <- attributable_pip(df_gene,combined_pip_col = "combined_pip")

DT::datatable(df_gene,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','top20 Genes sorted bu susiepip, 0 means pip=0; NA means no such gene in this tissue'),options = list(pageLength = 10) )

attributable_columns <- colnames(df_gene)[seq(4,12,by=2)]

plot_stacked_histogram(df_gene, "genename", "combined_pip", attributable_columns)

table_show <- ctwas_res[ctwas_res$genename %in% df_gene$genename,]
table_show <- rbind(table_show,ctwas_res[ctwas_res$type!="SNP" & ctwas_res$susie_pip > 0.8,])
table_show <- table_show[!duplicated(table_show$id),]
DT::datatable(table_show,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Detailed table for genes showed above'),options = list(pageLength = 5))

```


# OxHb

```{r fig.height=6, fig.width=12,message=FALSE, warning=FALSE}
outputdir <- "/project/xinhe/xsun/ctwas/2.TIBETN/output/multi/OxHb/"
outname <- "OxHb_ctwas"
gwas_n <- 966
weight <- "Heart_Left_Ventricle,Kidney_Cortex,Liver,Artery_Coronary,Lung"

load(paste0(outputdir,outname,"_z_snp.Rd"))
load(paste0(outputdir,outname,"_z_gene.Rd"))

ctwas_parameters <- ctwas:::ctwas_summarize_parameters(outputdir = outputdir,
                                               outname = outname,
                                               gwas_n = gwas_n,
                                               thin = thin)

ctwas_parameters$convergence_plot

ctwas_res <- process_data(weight = weight,outputdir = outputdir,outname = outname,z_snp = z_snp,z_gene = z_gene)
para <- process_ctwas_results(ctwas_parameters = ctwas_parameters,ctwas_res = ctwas_res)
DT::datatable(para,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Parameters'))

df_gene <- process_gene_data(ctwas_res = ctwas_res)

df_gene <- attributable_pip(df_gene,combined_pip_col = "combined_pip")

DT::datatable(df_gene,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','top20 Genes sorted bu susiepip, 0 means pip=0; NA means no such gene in this tissue'),options = list(pageLength = 10) )

attributable_columns <- colnames(df_gene)[seq(4,12,by=2)]

plot_stacked_histogram(df_gene, "genename", "combined_pip", attributable_columns)

table_show <- ctwas_res[ctwas_res$genename %in% df_gene$genename,]
table_show <- rbind(table_show,ctwas_res[ctwas_res$type!="SNP" & ctwas_res$susie_pip > 0.8,])
table_show <- table_show[!duplicated(table_show$id),]
DT::datatable(table_show,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Detailed table for genes showed above'),options = list(pageLength = 5))

```


# Pulse

```{r fig.height=6, fig.width=12,message=FALSE, warning=FALSE}
outputdir <- "/project/xinhe/xsun/ctwas/2.TIBETN/output/multi/Pulse/"
outname <- "Pulse_ctwas"
gwas_n <- 966
weight <- "Heart_Atrial_Appendage,Kidney_Cortex,Artery_Tibial,Whole_Blood,Liver"

load(paste0(outputdir,outname,"_z_snp.Rd"))
load(paste0(outputdir,outname,"_z_gene.Rd"))

ctwas_parameters <- ctwas:::ctwas_summarize_parameters(outputdir = outputdir,
                                               outname = outname,
                                               gwas_n = gwas_n,
                                               thin = thin)

ctwas_parameters$convergence_plot

ctwas_res <- process_data(weight = weight,outputdir = outputdir,outname = outname,z_snp = z_snp,z_gene = z_gene)
para <- process_ctwas_results(ctwas_parameters = ctwas_parameters,ctwas_res = ctwas_res)
DT::datatable(para,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Parameters'))

df_gene <- process_gene_data(ctwas_res = ctwas_res)

df_gene <- attributable_pip(df_gene,combined_pip_col = "combined_pip")

DT::datatable(df_gene,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','top20 Genes sorted bu susiepip, 0 means pip=0; NA means no such gene in this tissue'),options = list(pageLength = 10) )

attributable_columns <- colnames(df_gene)[seq(4,12,by=2)]

plot_stacked_histogram(df_gene, "genename", "combined_pip", attributable_columns)

table_show <- ctwas_res[ctwas_res$genename %in% df_gene$genename,]
table_show <- rbind(table_show,ctwas_res[ctwas_res$type!="SNP" & ctwas_res$susie_pip > 0.8,])
table_show <- table_show[!duplicated(table_show$id),]
DT::datatable(table_show,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Detailed table for genes showed above'),options = list(pageLength = 5))

```
