---
title: "Single tissue analysis (LD merge =T)"
author: "XSun"
date: "2023-12-14"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Data 

## Used in cTWAS analysis: 

- Traits (4): LB, MC, OxHb, Pulse

- Weights:Expression models from 9 tissues: Artery_Coronary, Artery_Tibial, Heart_Atrial_Appendage, Heart_Left_Ventricle, Kidney_Cortex, Liver, Lung, Pancreas, Whole_Blood. PredictDB, MASHR-based models.

- Weights for negtive control:Expression models from 6 tissues: Skin_Sun_Exposed_Lower_leg Nerve_Tibial Esophagus_Mucosa Breast_Mammary_Tissue Testis Brain_Cerebellum

- LD reference: UKBB


## Functions

```{r }
suppressMessages(library(readxl))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(forcats))
suppressMessages(library(gridExtra))
```


# Results

```{r warning=F, fig.height=10, fig.width=10}

dat <- read_xlsx("/project/xinhe/xsun/ctwas/2.TIBETN/results/sum.xlsx")

DT::datatable(dat,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary table'),options = list(pageLength = 10) )

dat<- dat %>% 
  distinct(traits, weights, .keep_all = TRUE)
dat <- dat[,1:7]

dat_sort <- dat %>%
  mutate(group_pve_gene = as.numeric(group_pve_gene)) %>%  # Convert to numeric if necessary
  group_by(traits) %>%
  arrange(desc(group_pve_gene), .by_group = TRUE) %>%  # Sort in descending order within each trait
  ungroup()

top3_per_trait <- dat_sort %>% 
  group_by(traits) %>% 
  top_n(3, attributable_pve_gene) %>% 
  ungroup()


traits <- c("LB","MC","OxHb","Pulse")
p <- list()
for (i in 1:length(traits)) {
  
  df_trait <- top3_per_trait[top3_per_trait$traits == traits[i],]
  df_trait$traits <- fct_inorder(df_trait$traits)
  df_trait$weights <- fct_inorder(df_trait$weights)
  df_trait$attributable_pve_gene <- as.numeric(df_trait$attributable_pve_gene)
  
  p[[i]] <- ggplot(df_trait , aes(x = weights, y = attributable_pve_gene, fill = weights)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~traits, scales = "free_x", nrow = 1) +  # Create clusters for each trait
    scale_fill_brewer(palette = "Set1") +  # Use a color palette
    labs(title = ,
         x = "Tissues",
         y = "Attributable PVE Gene") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    guides(fill = FALSE) +
    ylim(0, 0.15)
}


P <- grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]], nrow = 2, ncol = 2, top="Bar plot of the percent of gene PVE (attributed proportion) in the top 3 tissues per trait")

```


```{r warning=F, fig.height=10, fig.width=10}

dat <- read_xlsx("/project/xinhe/xsun/ctwas/2.TIBETN/results/sum.xlsx")

dat_all <- c()
p <- list()
for (i in 1:length(traits)) {
  
  dat_trait <- dat[dat$traits == traits[i],]
  dat_trait_processed <- dat_trait %>%
    mutate(group_pve_gene = as.numeric(group_pve_gene)) %>%  # Convert to numeric if necessary
    group_by(traits) %>%
    arrange(desc(group_pve_gene), .by_group = TRUE) %>%  # Sort in descending order within each trait
    ungroup()
  
  weights_top3 <- dat_trait_processed$weights[!duplicated(dat_trait$weights)][1:3]
  
  dat_trait_select <- dat_trait_processed[dat_trait_processed$weights %in% weights_top3,]
  # index_keep <- c(seq(1,nrow(dat_trait_select),by=5),seq(2,nrow(dat_trait_select),by=5),seq(3,nrow(dat_trait_select),by=5))
  # index_keep <- sort(index_keep)
  # dat_trait_select <- dat_trait_select[index_keep,]
  
  dat_trait_select$weights <- fct_inorder(dat_trait_select$weights)
  
  dat_trait_select <- dat_trait_select %>%
    mutate(genename = ifelse(duplicated(genename), 
                             paste0(genename, " "), 
                             genename))
  
  dat_trait_select$genename <- fct_inorder(dat_trait_select$genename)

  
  
  plot <- ggplot(dat_trait_select, aes(x = genename, y = susie_pip, fill = weights)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~weights, scales = "free_x", nrow = 1) +  # Create clusters for each trait
    scale_fill_brewer(palette = "Set1") +  # Use a color palette
    labs(title = traits[i],
         x = "Genes",
         y = "Susie PIP") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(fill = FALSE) +
    ylim(0, 0.45)
  p[[i]] <-plot 
  

}

P <- grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]], nrow = 2, ncol = 2, top="Bar plot for the top gene PIP in the top 3 tissues per trait")

```

# Results for negtive control

```{r warning=F, fig.height=10, fig.width=10}

dat <- read_xlsx("/project/xinhe/xsun/ctwas/2.TIBETN/results/sum_ng.xlsx")

DT::datatable(dat,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary table'),options = list(pageLength = 10) )

dat<- dat %>% 
  distinct(traits, weights, .keep_all = TRUE)
dat <- dat[,1:7]

dat_sort <- dat %>%
  mutate(group_pve_gene = as.numeric(group_pve_gene)) %>%  # Convert to numeric if necessary
  group_by(traits) %>%
  arrange(desc(group_pve_gene), .by_group = TRUE) %>%  # Sort in descending order within each trait
  ungroup()

top3_per_trait <- dat_sort %>% 
  group_by(traits) %>% 
  top_n(3, attributable_pve_gene) %>% 
  ungroup()


traits <- c("LB","MC","OxHb","Pulse")
p <- list()
for (i in 1:length(traits)) {
  
  df_trait <- top3_per_trait[top3_per_trait$traits == traits[i],]
  df_trait$traits <- fct_inorder(df_trait$traits)
  df_trait$weights <- fct_inorder(df_trait$weights)
  df_trait$attributable_pve_gene <- as.numeric(df_trait$attributable_pve_gene)
  
  p[[i]] <- ggplot(df_trait , aes(x = weights, y = attributable_pve_gene, fill = weights)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~traits, scales = "free_x", nrow = 1) +  # Create clusters for each trait
    scale_fill_brewer(palette = "Set1") +  # Use a color palette
    labs(title = ,
         x = "Tissues",
         y = "Attributable PVE Gene") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    guides(fill = FALSE) +
    ylim(0, 0.15)
}


P <- grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]], nrow = 2, ncol = 2, top="Bar plot of the percent of gene PVE (attributed proportion) in the top 3 tissues per trait")

```


```{r warning=F, fig.height=10, fig.width=10}

dat <- read_xlsx("/project/xinhe/xsun/ctwas/2.TIBETN/results/sum_ng.xlsx")

dat_all <- c()
p <- list()
for (i in 1:length(traits)) {
  
  dat_trait <- dat[dat$traits == traits[i],]
  dat_trait_processed <- dat_trait %>%
    mutate(group_pve_gene = as.numeric(group_pve_gene)) %>%  # Convert to numeric if necessary
    group_by(traits) %>%
    arrange(desc(group_pve_gene), .by_group = TRUE) %>%  # Sort in descending order within each trait
    ungroup()
  
  weights_top3 <- dat_trait_processed$weights[!duplicated(dat_trait$weights)][1:3]
  
  dat_trait_select <- dat_trait_processed[dat_trait_processed$weights %in% weights_top3,]
  # index_keep <- c(seq(1,nrow(dat_trait_select),by=5),seq(2,nrow(dat_trait_select),by=5),seq(3,nrow(dat_trait_select),by=5))
  # index_keep <- sort(index_keep)
  # dat_trait_select <- dat_trait_select[index_keep,]
  
  dat_trait_select$weights <- fct_inorder(dat_trait_select$weights)
  
  dat_trait_select <- dat_trait_select %>%
    mutate(genename = ifelse(duplicated(genename), 
                             paste0(genename, " "), 
                             genename))
  
  dat_trait_select$genename <- fct_inorder(dat_trait_select$genename)

  
  
  plot <- ggplot(dat_trait_select, aes(x = genename, y = susie_pip, fill = weights)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~weights, scales = "free_x", nrow = 1) +  # Create clusters for each trait
    scale_fill_brewer(palette = "Set1") +  # Use a color palette
    labs(title = traits[i],
         x = "Genes",
         y = "Susie PIP") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(fill = FALSE) +
    ylim(0, 0.45)
  p[[i]] <-plot 
  

}

P <- grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]], nrow = 2, ncol = 2, top="Bar plot for the top gene PIP in the top 3 tissues per trait")

```

